---
title: "Homework Unit 3: KNN"
author: "Your name here"
date: "`r lubridate::today()`"
format: 
  html: 
    embed-resources: true
    toc: true 
    toc_depth: 4
editor_options: 
  chunk_output_type: console
editor: 
  markdown: 
    wrap: 72
---

## General Instructions

Welcome to the KNN portion of your Unit_3 homework! Please enter your
first and last name at the top of this document where it says "Your name
here".

This homework aligns with **Unit 3 Exploratory Introduction to
Regression Models** in the course web book, where you can download all
materials required for this assignment:

**Two Markdown files (`hw_unit_3_lm.qmd` and `hw_unit_3_knn.qmd`):** We
will have separate files for fitting lm and knn models for this
homework. The order you complete these in does not matter; both contain
all instructions needed for the assignment.

-   **`ames_train_cln.csv`**: The `.csv` file containing the cleaned
    training data to be used for this assignment.

-   **`ames_val_cln.csv`**: The `.csv` file containing the cleaned
    validation data to be used for this assignment.

-   **`ames_test_cln.csv`**: The `.csv` file containing the cleaned test
    data to be used for this assignment. `sale_price` has been removed
    from this file.

-   **`ames_data_dictionary.pdf`:** The data dictionary describing
    variables contained in the data set.

------------------------------------------------------------------------

The goal of this assignment is to create the best regression model you
can to predict housing sale prices in the Ames data set. To build this
model, you will have all variables used by John in the unit 2 & 3 web
book to demo EDA and Regression Models, as well as all variables that
you performed EDA on yourself in the unit 2 homework assignment.

A very nice graduate student combined the EDA cleaning steps from the
assignment and web book data to provide you with three data files
(training, validation, and test) containing all cleaned variables that
you need for this assignment.

Using `ames_train_cln` as your starting point, your job is to use simple
feature engineering techniques (e.g. selecting among variables,
transforming variables, coding categorical variables, adding
interactions) to iteratively train and compare multiple `lm` and `knn`
models. **This script is for training `lm` models.** You do not need to
perform any additional cleaning EDA, but you may continue to add on to
your homework 2 EDA modeling script as you consider different feature
engineering approaches. You do not need to turn in any additional
modeling EDA code you complete.

At the end of this assignment, you will select your best performing
model across all `lm` and `knn` models you fit. You will use
`ames_test_cln` *only once* at the end of the assignment to generate
your best model's predictions of `sale_price` in the held out test set.
The TAs will assess everyone's predictions on the held-out set to
determine the best fitting model in the class.

**The winner gets a *FREE LUNCH* with John! Good luck!** (＾▽＾)

Lets get started!

## Set up

### Handle conflicts

### Load required packages and source

Add packages (only what you need!)

```{r}
library(tidymodels)
library(tidyverse)
options(conflicts.policy = "depends.ok")
```

Source any function scripts (John's are already sourced for you, buta
you can change these or add your own function scripts too!)

```{r}
#| message: false


devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/fun_plots.R?raw=true")
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/fun_eda.R?raw=true")
```

### Specify other global settings

```{r}
#| include: false

theme_set(theme_classic())
options(tibble.width = Inf, dplyr.print_max=Inf)
```

### Paths

```{r}
setwd("C:/Users/gujia/Desktop/github/book_iaml/application_assignments/unit_03")
```

### Load data

Load the cleaned training, validation, and test data files. **Please
load your data using the filenames provided above**

Use `here::here()` and relative path for your data. Make sure your iaml
project is open

```{r}
data_trn <- read_csv("ames_train_cln.csv")
data_val <- read_csv("ames_val_cln.csv")
data_test <- read_csv("ames_test_cln.csv")
```

set appropriate variable classes (remember character classes should be
factors!)

```{r}
data_trn |>
  mutate(across(where(is.character),factor))|>
  glimpse()
```

### Create tracking tibble

Create a tibble to track the validation errors across various model
configurations.

```{r}
error_val <- tibble(model = character(), rmse_val = numeric())
```

## KNN Model building instructions

You may consider as many KNN model configurations as you wish, but there
are a few rules:

-   All models must eventually predict **raw** sale_price (advanced
    hint: you may want to transform the outcome, but then if you do,
    make sure you transform the predictions back to raw \$ when you
    submit model predictions; and if that doesnt make sense, don't worry
    about it. Just work with raw dollars!)

-   You must build a minimum of 3 models

-   At least one model must include two or more range corrected numeric
    predictors

-   At least one model must include a categorical predictor with more
    than two levels

-   At least one model must include a categorical variable with levels
    that have been modified during feature engineering (NOT including
    dummy coding)

-   At least one model should vary values of *k*

Your models may include as many other predictors as you would like (e.g.
your model satisfying the 2 numeric predictors requirement may also
include categorical variables, more than 2 numeric variables, etc.) so
be as creative as you'd like!

## KNN 1

List out which required model components you are including for your
first model (delete ones that do not apply): *2+ numeric predictors*
*Categorical \> 2 levels* *Modified categorical* *Varied `k`*

### Set up recipe

Informed by EDA from your modeling EDA script, create a recipe for your
first model.

```{r}
rec <- recipe(sale_price ~ gr_liv_area + lot_area + total_bsmt_sf, data = data_trn)|>
  step_scale(all_numeric_predictors())
```

### Training feature matrix

Use the `prep()` function to prep your recipe

```{r}
rec_prep <- rec|>
  prep(training = data_trn)
```

Use the `bake()` function to generate the feature matrix of the training
data.

```{r}
feat_trn_1 <- rec_prep |>
  bake(NULL)
```

### Fit your model

Fit a `knn` model predicting raw `sale_price` from your training feature
matrix.

```{r}
fit_5nn_3nn <-
  nearest_neighbor()|>
  set_engine("kknn")|>
  set_mode("regression")|>
  fit(sale_price ~ gr_liv_area + lot_area + total_bsmt_sf, data = feat_trn_1)
```

### Validation feature matrix

Use `bake()` to generate the feature matrix of the validation data that
we will use to assess your model.

```{r}
feat_val_1 <- rec_prep |>
  bake(data_val)
```

### Assess your model

Use `rmse_vec()` to calculate the validation error (RMSE) of your model.
Add this value to your validation error tracking tibble.

```{r}
error_val <- bind_rows(error_val, 
                        tibble(model = "3 numeric predictor 3nn", 
                               rmse_val = rmse_vec(feat_val_1$sale_price, 
                                                   predict(fit_5nn_3nn, feat_val_1)$.pred)))

error_val
```

### Visualize performance

Visualize the relationship between raw and predicted sale price in your
validation set using the `plot_truth()` function.

```{r}
plot_truth <- function(truth, estimate) {
  ggplot(mapping = aes(x = truth, y = estimate)) + 
    geom_abline(lty = 2, color = "red") + 
    geom_point(alpha = 0.5) + 
    labs(y = "predicted outcome", x = "outcome") +
    coord_obs_pred()   # scale axes uniformly (from tune package)
}
```

```{r}
plot_truth(truth = feat_val_1$sale_price, 
           estimate = predict(fit_5nn_3nn, feat_val_1)$.pred)
```

## KNN 2

List out which required model components you are including for your next
model (delete ones that do not apply): *2+ numeric predictors*
*Categorical \> 2 levels* *Modified categorical* *Varied `k`*

### Set up recipe

Informed by EDA from your modeling EDA script, create a recipe for your
next model.

```{r}
rec_2 <- recipe(sale_price ~ ms_zoning, data = data_trn)|>
  
  step_mutate(ms_zoning = fct_collapse(ms_zoning, "residential" = c("rh","rm","rl"), "commercial" = c("a", "c", "i"), "floating" = "fv"))|> step_dummy(ms_zoning)|> 
  step_scale(all_numeric_predictors())|>
  step_center(all_numeric_predictors())
```

### Training feature matrix

Use the `prep()` function to prep your recipe

```{r}
rec_prep_2 <- rec_2 |> prep(data_trn)
```

Use the `bake()` function to generate the feature matrix of the training
data.

```{r}
feat_trn_2 <- rec_prep_2 |> bake(NULL)
```

### Fit your model

Fit a `knn` model predicting raw `sale_price` from your training feature
matrix.

```{r}
fit_20nn_1cat <-
  nearest_neighbor(neighbors = 20)|>
  set_engine("kknn")|>
  set_mode("regression")|>
  fit(sale_price ~ ., data = feat_trn_2)
```

### Validation feature matrix

Use `bake()` to generate the feature matrix of the validation data that
we will use to assess your model.

```{r}
feat_val_2 <- rec_prep_2 |> bake(data_val)
```

### Assess your model

Use the `rmse_vec()` function to calculate the validation error (RMSE)
of your model. Add this value to your validation error tracking tibble.

```{r}
error_val <- bind_rows(error_val, tibble(model = "1 categorical predictor 20nn", rmse_val = rmse_vec(feat_val_2$sale_price, predict(fit_20nn_1cat, feat_val_2)$.pred))) 
error_val
```

### Visualize performance

Visualize the relationship between raw and predicted `sale_price` in
your validation set using `plot_truth()`.

```{r}
plot_truth(truth = feat_val_2$sale_price, estimate = predict(fit_20nn_1cat, feat_val_2)$.pred)
```

## KNN 3

List out which required model components you are including for your next
model (delete ones that do not apply): *2+ numeric predictors*
*Categorical \> 2 levels* *Modified categorical* *Varied `k`*

### Set up recipe

Informed by EDA from your modeling EDA script, create a recipe for your
next model.

```{r}
rec_3 <- recipe(sale_price ~ overall_qual, data = data_trn)|> step_mutate(overall_qual = as.numeric(overall_qual))|> step_scale(all_numeric_predictors())
```

### Training feature matrix

Use the `prep()` function to prep your recipe

```{r}
rec_prep_3 <- rec_3|> prep(data_trn)
```

Use the `bake()` function to generate the feature matrix of the training
data.

```{r}
feat_trn_3 <- rec_prep_3|> bake(NULL)
```

### Fit your model

Fit a `knn` model predicting raw `sale_price` from your training feature
matrix.

```{r}
fit_5nn_1ord <- nearest_neighbor() |> set_engine("kknn")|> set_mode("regression")|> fit(sale_price ~ ., data = feat_trn_3)
```

### Validation feature matrix

Use `bake()` to generate the feature matrix of the validation data that
we will use to assess your model.

```{r}
feat_val_3 <- rec_prep_3 |> bake(data_val)
```

### Assess your model

Use `rmse_vec()` to calculate the validation error (RMSE) of your model.
Add this value to your validation error tracking tibble.

```{r}
error_val <- bind_rows(error_val, tibble(model = "1 ordinal predictor 5nn", rmse_val = rmse_vec(feat_val_3$sale_price, predict(fit_5nn_1ord, feat_val_3)$.pred))) 
error_val
```

### Visualize performance

Visualize the relationship between raw and predicted `sale_price` in
your validation set using `plot_truth()`.

```{r}
plot_truth(truth = feat_val_3$sale_price, estimate = predict(fit_5nn_1ord, feat_val_3)$.pred)
```

## 4 Additional configurations

Create as many code chunks as you would like below to test additional
linear model configurations. Record your models' performance in your
validation error tracking tibble.

```{r}
```

## Predictions

This section is for generating predictions for your best model in the
held out test set. **You should only generate predictions for ONE model
out of all `lm` and `knn` models you fit across both homework `.Rmd`
files.** The TAs will use these predictions to generate your ONE
estimate of model performance in the held out data.

### Assign best model

If your **best** model is a `knn` from this script, edit the following
objects in the code chunk below:

-   Assign the model object from your best model to `best_model`
-   Copy and paste the recipe for your best model (created from the
    training data) to assign to `best_rec` below.
-   Type your last name between the quotation marks of `last_name` (e.g.
    "santana") for file naming.

If your best model is a `lm`, don't edit anything in the next two code
chunks and skip to the final section to save, render, etc.

```{r}
best_model <- NULL
best_rec <- NA
last_name <- ""
```

### Generate test predictions

Run this code chunk to save out your best model's predictions of raw
`sale_price` in the held-out test set. Look over your predictions to
confirm your model generated valid predictions for each test
observation. (If you left `best_model` as NA above, no file will be
saved out).

ADVANCED NOTE: You *could* improve this code below to train an even
better model. If you know what could be improved, do it. If not, do not
worry. This will work well enough (but maybe not for a free lunch!!).

ADVANCED NOTE 2: If your model used a transformation of `sale_price` you
will also need to update this code.

```{pred-test}
if (!is.null(best_model)){
  
 # Make test set features with the best recipe
  rec_prep <- best_rec |> 
    prep(training = data_trn)
  
  feat_test <- rec_prep |> 
    bake(new_data = data_test)
  
 # Generate predictions made by the best model
  test_preds <- data_test |> 
    mutate(sale_price = predict(best_model, feat_test)$.pred) |> 
    select(pid, sale_price) #pid is the id variable to match predictions
  
  # Save predictions as a csv file with your last name in the file name
   write.csv(test_preds, here::here(path_data,str_c("test_preds_", 
                                                   last_name,
                                                   ".csv")))
}
```

## Save, knit, upload, and remember

### Save

Save this `.qmd` file with your last name at the end (e.g.
`hw_unit_3_knn_santana.qmd`)

### Knit

Knit the file to html and upload your knit html file to Canvas.

### Upload

Upload your saved `test_preds_(lastname).csv` (if you created it in this
script) to Canvas. We will assess everyones predictions in the test set.
**The winner gets a high five from John in front of the whole class as
well as a free lunch!!!**

### Remember...there's more

Don't forget to complete the `lm` portion of this assignment
(`hw_unit_3_lm.qmd`) as well.

**\~ \~ \~ You are indeed a machine learning star - and we are proud of
you for it \~ \~ \~**
